---
title: "Craft Beer Breweries"
author: "Pearly Paul and Kevin Albright"
date: "6/21/2021"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
  html_document:
    df_print: paged
---
## Introduction 
​
In the below report, we have been provided two data sets for various `beers` and the `breweries` that produce them in the US. 
Our team had to do an exploratory Data Analysis on the data provided to come up with useful insights.
​
We were provided with two separate data sets, one with details of the beers and and the other with the details of the breweries around the USA. To analyze this 
​
​
## Reproducible Research in R
​
Please begin by reading the the README.md.All the details on the structure of project has been documented.
​
### Loading Packages for Analysis
Below are the packages required for the analysis of the Craft Beer and Brewery study.
```{r setup, include=FALSE}
#install.packages("tidyverse")
#install.packages("usdata")
#install.packages("ggthemes")
#install.packages("viridis")
#install.packages("class")
#install.packages("caret")
#install.packages("prettydoc")
#install.packages("kable_extra")
#
library(class)
library(kableExtra)
library(caret)
library(e1071)
library(viridis)
library(ggthemes)
library(usdata)
library(tidyverse)
library(dplyr)
library(sqldf)
library(prettydoc)
knitr::opts_chunk$set(echo = TRUE)
```

### Loading the Datasets for Analysis 
​
We were provided with 2 data sets (`beers` and `breweries`). The `breweries` dataset contains the count of breweries from 50 states of the US and the `beer` data set contains the details of the beers (Alcohol by Volume Content and International Bitterness unit)
​
`Note : Beers dataset contains a list of 2410 US craft beers and Breweries dataset contains 558 US breweries.`
​
***Region Data : This is the data set we created for analysis group the 50 states in US into 4 regions North Central,North East , South and West***



```{r load and clean data}

#Read in Core Data
Beers = read.csv("/Users/Kevin/Desktop/School/Doing Data Science/Project 1/Beers.csv", header = TRUE)
Breweries = read.csv("/Users/Kevin/Desktop/School/Doing Data Science/Project 1/Breweries.csv", header = TRUE)

#Bring in Region Data
RegionData = data.frame(State = state.abb, Region = state.region)
Breweries = left_join(Breweries,RegionData,by = "State")

#Handle NA and missing data by replacing with Regional Means
Breweries = sqldf('
      select
      "Brew_ID","Name","City","State",
      
      case when "State"  = "DC" then "South" else "Region" end as Region
      
      from
      Breweries
                 ')

Distilled_Data = merge(Beers,Breweries, by.x = "Brewery_id", by.y = "Brew_ID")
#Distilled_Data = left_join(Beers,Distilled_Data,by = "Beer_ID")

#Create table for IBU and ABV means by State and Region - excluding missing data 
State_Means = 
Distilled_Data[!is.na(Distilled_Data$ABV) & !is.na(Distilled_Data$IBU),] %>% 
  group_by(State) %>%
  summarize(Mean_IBU_by_State = mean(IBU),
            Mean_ABV_by_State = mean(ABV))
Regional_Means = 
  Distilled_Data[!is.na(Distilled_Data$ABV) & !is.na(Distilled_Data$IBU),] %>% 
  group_by(Region) %>%
  summarize(Mean_IBU_by_Region = mean(IBU),
            Mean_ABV_by_Region = mean(ABV))

Distilled_Data = left_join(Distilled_Data,State_Means,by = "State")
Distilled_Data = left_join(Distilled_Data,Regional_Means,by = "Region")


```
## 3. Research Questions
​
We are now going to analyze the research questions for EDA 
​
​
### 1. How many breweries are present in each state?
​
Here we used the breweries dataset to count the number of breweries per state in the US.
​
**Output : We were able to find Colorado and California to have the highest number of Craft breweries in US.**
```{r count of Breweries by State}

kable(count(Breweries,State)) 

```

```{r Brewery Count by State,fig.width=10,fig.height=7}
#Brewery Count by State
Breweries %>% 
  group_by(State,Region) %>%
  summarize(Count = n()) %>%
  ggplot(aes(x = reorder(State,-Count), y = Count,fill=Region)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label=Count),vjust=1) + 
  ggtitle("Brewery Count by State",) + xlab("State") + ylab("Count") + 
  facet_wrap(~Region, scales="free") + 
  theme(plot.title=element_text(hjust=.5)) 
```
## 2. Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.
​
Here we merged the both the data sets into a single data set `Distilled_Data`. By which we can find the beers made in by each state in United States.
​
The  Brewery_id column  name in the Beers data was changed to Brew_id for merging both data sets
#Address missing data and Plot Median ABV and Bitterness
```{r Distilled Data head}
kable(head(Distilled_Data))
kable(tail(Distilled_Data))
write.csv(Distilled_Data,"/Users/Kevin/Desktop/School/Doing Data Science/Project 1/Distilled_Data.csv")
write.csv(RegionData,"/Users/Kevin/Desktop/School/Doing Data Science/Project 1/Region_Data.csv")
```

## 3.Address the missing values in each column.
​
To proceed with data analysis we first inspected both the datasets for any missing value.
​
**1.Beer Data set :** 
​
  + International Bitterness Unit : Out of the 2410 beers - 1405 beers were missing IBU value
  + Alcohol By Volume : Out of the 2410 beers - 62 beers were missing ABV content value 
  
**2.Breweries Data set :** 
  + No missing data 
  
To deal with this anomaly and unbiased data analysis we populated the missing values in the Beer Data. We populated the missing values with the mean of respective Region/State.
```{r handle missing data}
Distilled_Data2 = Distilled_Data

Distilled_Data2 = 
Distilled_Data2 %>%
  mutate(ABV = coalesce(Distilled_Data2$ABV,Distilled_Data2$Mean_ABV_by_State),
         IBU = coalesce(coalesce(Distilled_Data2$IBU,Distilled_Data2$Mean_IBU_by_State),Distilled_Data2$Mean_IBU_by_Region))


```
## 4.Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
​
***Calculation of the median Alcohol content of each state grouped by region***
​
​
### +4a. Median Alcohol Content - After plotting we discovered that the state with the highest median ABV was Kentucky with an ABV of .064 hailing from the Southern Region.  The median bitterness was highest in West Virginia again from the Southern Region.
```{r median ABV and IBU,fig.width=10,fig.height=7}
#Median ABV                        
Distilled_Data2 %>%
  group_by(State,Region) %>%
  summarize(
    Median_Alcohol_Content = round(median(ABV,na.rm=TRUE),digits=5), 
    Median_Bitterness = median(IBU,na.rm=TRUE)
  ) %>%
  ggplot(aes(reorder(x = State,Median_Alcohol_Content), y = Median_Alcohol_Content,fill=Region)) + geom_bar(stat="identity") + 
  xlab("State") + ylab("ABV") + ggtitle("Median ABV by State") + facet_wrap(~Region, scales="free") + 
  geom_text(aes(label=round(Median_Alcohol_Content,digits=6)),hjust=1) + 
  coord_flip()
#Median Bitterness
Distilled_Data2 %>%
  group_by(State,Region) %>%
  summarize(
    Median_Bitterness = median(IBU,na.rm=TRUE),
    Median_Alcohol_Content = median(ABV,na.rm=TRUE) 
  ) %>%
  ggplot(aes(reorder(x = State,Median_Bitterness), y = Median_Bitterness,fill=Region)) + geom_bar(stat="identity") + 
  xlab("State") + ylab("Bitterness") + ggtitle("Median Bitterness by State") + coord_flip() +
  geom_text(aes(label=round(Median_Bitterness,digits=2)),hjust=1) + 
  facet_wrap(~Region, scales="free") 
```
## 5.Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
​
### 5a. Most Alcoholic Beer in US
​
### Observation : Upon analysis the beer with maximum alcohol is made in the `Lee Hill Series Vol.5`  made in Colorado West

### 5b. Most Bitter Beer in US

### Observation : Upon analysis the beer with maximum bitter content is  `Bitter Bitch Imperial IPA`  made in Oregano in West Region
```{r pressure, echo=TRUE}
Top_ABV = Distilled_Data2[order(-Distilled_Data2$ABV),] 
kable(head(Top_ABV,1)) 
```
```{r Most Bitter Beer}
Most_Bitter = Distilled_Data2[order(-Distilled_Data2$IBU),] 
kable(head(Most_Bitter,1))
```
## 6.   Comment on the summary statistics and distribution of the ABV variable.
### Observation :  The chart shows a right tailed distribution with the majority of alcohol content hovering between .05 and .06
```{r ABV Summary Stats,fig.width=10,fig.height=7}
summary(Distilled_Data2$ABV)
boxplot(Distilled_Data2$ABV,main="ABV")
```
### 7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  
## Observation :  We observed a positive linear correlation between ABV and IBU.  Overall, as the bitterness increased so did the alcohol content.
```{r Relationship Between ABV and IBU,fig.width=10,fig.height=7}
Distilled_Data2 %>% ggplot(aes(x=IBU, y=ABV)) + geom_point(color="red") + geom_smooth(method="lm") + 
  ggtitle("Relationship Between ABV and IBU")
```

### 8. Use KNN to differentiate the beer from Ale or IPA Beer
## Observation :  First we had to filter the data set to only ALE and IPA beers then plot the relationship by region.  
```{r KNN,fig.width=10,fig.height=7}
#Relationship Between IBU and ABV by Type
IPA_ALE = sqldf('
  select
  "Name","Style" as Style,"ABV","IBU","State","Region","City",
  case when "Style" like "%ALE%" then "ALE"
       when "Style" like "%IPA%" then "IPA" end as Type
 from
  Distilled_Data2
 where
  "Style" like "%ALE%" OR "Style" like "%IPA%"
 ')

IPA_ALE %>% ggplot(aes(x=IBU, y=ABV,color=Type)) + geom_point() + geom_smooth(method="lm") + 
  ggtitle("Relationship Between ABV and IBU") + 
  facet_wrap(~Region,scales="free")
```
### 8b. Use KNN to differentiate the beer from ALE or IPA Beer
## Observation : Found optimal K in order to produce the most accurate KNN model.  When k = 5 we were able to formulate an 84% accuracy using the variables ABV and IBU
```{r Implement KNN,fig.width=10,fig.height=7}
splitPerc = .70
trainBeer = sample(1:dim(IPA_ALE)[1],round(splitPerc * dim(IPA_ALE)[1]))
train = IPA_ALE[trainBeer,]
test = IPA_ALE[-trainBeer,]

accs = data.frame(accuracy = numeric(90), k = numeric(90))

for(i in 1:90)
{
  classifications = knn(train[,c(3,4)],test[,c(3,4)],train$Type, prob = TRUE, k = i)
  table(test$Type,classifications)
  CM = confusionMatrix(table(test$Type,classifications))
  accs$accuracy[i] = CM$overall[1]
  accs$k[i] = i
}
plot(accs$k,accs$accuracy, type = "l", xlab = "k") 
abline(v=accs$k[which.max(accs$accuracy)], col="red")
accs$k[which.max(accs$accuracy)]

splitPerc = .70
trainIndices = sample(1:dim(IPA_ALE)[1],round(splitPerc * dim(IPA_ALE)[1]))
train = IPA_ALE[trainIndices,]
test = IPA_ALE[-trainIndices,]

classification = knn(IPA_ALE[,c(3,4)],IPA_ALE[,c(3,4)],IPA_ALE$Type,prob = TRUE, k = 5)
table(classification,IPA_ALE$Type)
confusionMatrix(table(classification,IPA_ALE$Type))
```
#Additional Insight on Craft beer's and breweries
## Observation : We found that the southern and western regions consume alcohol with higher bitterness and the northeast and northern regions prefer a less alcoholic content in their beer.  The geographipal map of US shows the state groupings by regions.  
```{r GeoMap,fig.width=10,fig.height=7 }
map_data = map_data('state')
map_data = map_data %>% mutate(State = state2abbr(map_data$region))

Final_Breweries = left_join(Breweries,map_data, by = 'State')
Count_Breweries = Breweries %>%
  group_by(State) %>%
  summarize(Tally = n())

New_Breweries = left_join(Final_Breweries,Count_Breweries,by = 'State')
p0 <- ggplot(data = New_Breweries,
             mapping = aes(x = long, y = lat,
                           group = group,fill=Tally))
p1 <- p0 + geom_polygon(color = "black", size = 0.1) + 
  theme_map() +
  scale_fill_gradient2(low = "green",
                       mid = "yellow",
                       high = "red") +
  ggtitle("Breweries By Region") + 
  facet_grid(~Region, scales="free")
p1 
```